---
title: "Karthikeyan Chellamuthu_10.2.2"
author: "Karthikeyan Chellamuthu"
date: "2/19/2022"
---

Fit a logistic regression model to the binary-classifier-data.csv dataset from the previous assignment.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "D:/BU/DSC 520 T302-2221 winter 2021-22/GIT-Hub/dsc520-master/data")
```

```{r echo = FALSE, message=FALSE}
# Setup Work Directory where document is present
library(class)
library(ggplot2)
library(broom)
library(QuantPsyc)
library(car)
library(caret)
library(e1071)
library(caTools)
options(scipen = 999)
setwd("D:/BU/DSC 520 T302-2221 winter 2021-22/GIT-Hub/dsc520-master/data")
Binary_classifier <- read.csv("binary-classifier-data.csv")
```

## 10.2.2.a What is the accuracy of the logistic regression classifier?

```{r echo=FALSE}
train_index=createDataPartition(Binary_classifier$label, p=0.6)$Resample1
## Create trainig and testing set
training=Binary_classifier[train_index,] 
test=Binary_classifier[-train_index,]
```

```{r}
glm1 <- glm(label ~ x + y,family = binomial(),data = training)

tidy(glm1)
```

```{r echo=FALSE}
fittedSample=predict(glm1, newdata = test, type="response")
ConfMat=table(Actual = test$label,Predicted = fittedSample < 0.5)
Accuracy=sum(diag(ConfMat))/sum(ConfMat)
```

Accuracy of the model : **`r round(Accuracy * 100, digit = 2)`%**

## 10.2.2.b How does the accuracy of the logistic regression classifier compare to the nearest neighbors algorithm?

```{r echo=FALSE}
control <- trainControl(method="repeatedcv",repeats = 10)
KNNFind <- train(label ~ ., data = training, method = "knn", trControl = control, preProcess = c("center","scale"),tuneLength = 60)
KNNFind
plot(KNNFind)
```

```{r}
kNNM <- knn(train = training, test = test, cl = training$label, k = 7)
confusionMatrix(table(kNNM, test$label))
```
Accuracy of the kNN model : **97%**

## 10.2.2.c Why is the accuracy of the logistic regression classifier different from that of the nearest neighbors?

+ KNN is lazy execution, which means it fits and predicts at the time of prediction. KNN is better than logistic regression when the data contains high SNR
+ KNN is a completely non-parametric approach: No assumptions are made about the shape of the decision boundary. 
+ KNN supports non-linear solutions where LR supports only linear solutions.
+ KNN does not tell us which predictors are important as we don’t get a table of coefficients with p-values.