---
title: "Karthikeyan Chellamuthu 11.2.2 Exercise  Clustering "
author: "Karthikeyan Chellamuthu"
date: "03/05/2022"
output: PDF
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE, message=FALSE}
# Setup Work Directory where document is present
library(class)
library(ggplot2)
library(caret)
library(fpc)
options(scipen = 999)
setwd("D:/BU/DSC 520 T302-2221 winter 2021-22/GIT-Hub/dsc520-master/data")
cluster_data <- read.csv("clustering-data.csv")
```

## Structure of Clustering data
```{r echo=FALSE}
str(cluster_data)
```

## 11.2.a Plot the dataset using a scatter plot.
```{r echo=FALSE}
plot1 <- ggplot(cluster_data, aes(x, y)) +
  geom_point(alpha = 0.4) +
  labs(title="Scatter plot:  Clustering data") +
  theme_bw()

plot1
```

## 11.2.b Fit the dataset using the k-means algorithm from  k=2 to k=12. Create a scatter plot of the resultant clusters for each value of k.
```{r echo=FALSE}
kMeansCluster <- list()   ### empty list

for (i in 2:12){ 
  kMeansCluster[[i]] <- kmeans(cluster_data, centers = i)
  kplot <- ggplot(cluster_data, aes(x, y)) + 
  geom_point(aes(col=factor(kMeansCluster[[i]]$cluster)), alpha = 0.4) + 
  ggtitle(paste("kMeans Cluster Plot:  Clustering data - ",i," centroids"))
  
  print(kplot)
}
```

## 11.2.c Calculate this average distance from the center of each cluster for each value of k and plot it as a line chart where k is the x-axis and the average distance is the y-axis.

```{r echo=FALSE}

withinSqrtSum <- (nrow(cluster_data)-1)*sum(apply(cluster_data,2,var))
set.seed(1234)
for (i in 2:12)
  withinSqrtSum[i] <- sum(kMeansCluster[[i]]$withinss)

plot(1:12, withinSqrtSum, type="b", xlab = "K Value - Number of Clusters", ylab = "Within sum of squares", main = "Within groups sum of squares")

```

## 11.2.d One way of determining the “right” number of clusters is to look at the graph of k versus average distance and finding the “elbow point”. Looking at the graph you generated in the previous example, what is the elbow point for this dataset?


For k=5, between_ss/total_ss ratio tends to change slowly and remain less changing as compared to other k’s. Hence data k=5 should be a good choice for number of clusters.
 